\chapter{Acknowledgements}

I am deeply grateful to my supervisor, dr hab. Bartłomiej Gardas, whose unwavering support and insightful guidance have been instrumental in shaping this doctoral thesis. His expertise, encouragement, and mentorship have been invaluable, and I am truly fortunate to have had the opportunity to work under his supervision.

I would also like to express my sincere appreciation to my co-supervisor, dr hab. inż. Łukasz Pawela, for his constructive feedback, especially in the field of software engineering. His expertise and willingness to share knowledge have significantly enriched the quality of this thesis.

I would also like to express my gratitude to all my colleagues from the Institute who contributed to this thesis via many fruitful conversations I had with them. In particular, I would like to thank Krzysztof Domino for sharing his knowledge and expertise in the railway dispatching field.

Additionally, I extend my deepest gratitude to my friends, Alexander Juda and Michał Stęchły, for taking the time to read parts of this thesis. Their valuable feedback contributed greatly to improving the readability and overall quality of this work.

This project was partially supported by the National Science Center (NCN), Poland, under Projects: Sonata Bis 10, No. 2020/38/E/ST3/00269
and the National Centre for Research and Development (NCBR), Poland, under Project No. POIR.01.01.01-00-0061/2. I would also like to thank The Quantum Data Center Corporation for providing me with access to several GPUs used for benchmarks presented in this thesis.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
